---
title: "R Notebook"
output: rmarkdown::github_document
---

Required libraries

```{r}

library(readxl)
library(readr)
library(knitr)
library(tibble)
library(tidyr)
library(dplyr)

```



Tidy both tables

```{r}
# Read Table 1, unit is Decare (1 Decare = 1k m^2)
sown.area <- read_excel("data.xls", range = "R6C1:R25C7")
t1.col.names <- read_excel("data.xls", range = "R3C2:R4C7")
colnames(sown.area) <- c("Year",t1.col.names[1,])

# Read Table 2, unit is Tonne
production <- read_excel("data.xls", range = "R27C1:R46C7")
t2.col.names <- read_excel("data.xls", range = "R3C2:R4C7")
colnames(production) <- c("Year",t1.col.names[1,])

# Tidy Table 1
tidy.sown.area <- sown.area %>% gather(Type, Area, `Wheat`:`Sugar beets`)
tidy.sown.area


# Tidy Table 2

tidy.production <- production %>% gather(Type, Tonnes, Wheat:`Sugar beets`)
tidy.production
```



Using `merge` merge the to tables into one. Table `merged` has 4 columns: Year, Type, Area, Tonnes

```{r}
merged <- merge(tidy.sown.area, tidy.production)
merged
```



Add column `efficiency` to `merged`. `efficiency` is the amount of production per decare

```{r}
merged[,5] <- c(merged$Tonnes / merged$Area)
colnames(merged)[5] <- c("Efficiency")
merged
```



Scale `efficiency` using *normalization*

```{r}
rng <- merged[,5, drop = F] %>% apply(2, range)
rng <- rbind(rng, rng[2,] - rng[1,] )
rownames(rng) <- c("Min", "Max", "Range")
rng
```


```{r}
scaled <- cbind(
  merged %>% select(-Efficiency),
  merged %>% select(Efficiency) %>% scale(center = rng["Min",], scale = rng["Range", ])
)
scaled
# merged[,5, drop=F] %>% apply(2, range)
```



Discretize `efficiency` into 3 bins
1. low, medium, high

```{r}
discretized.range <- cut(merged$Efficiency, 3, labels = c("Low", "Medium", "High"))
range.table <- table(discretized.range)
barplot(range.table)
range.report <- tibble(Range = names(range.table), Count = range.table)
range.report
```

2. at breakpoints: (-0.01, 0.04, 0.10, 1.01)

```{r}
discretized.break <- cut(merged$Efficiency, breaks = c(-0.01, 0.04, 0.10, 1.01))
break.table <- table(discretized.break)
barplot(break.table)
break.report <- tibble(Break = names(break.table), Count = break.table)
break.report
```



Display a table to show how many times each product type achieved each efficiency level
##             
##              low medium high
##   Barley      18      0    0
##   Cotton       4     14    0
##   Maize        0     11    7
##   SugarBeets   0      0   18
##   Sunflower   18      0    0
##   Wheat       18      0    0

```{r}
# type.efficiency.table <- tibble(Type = colnames(sown.area)[-1])
label.appended <- merged %>% cbind(discretized.range)
colnames(label.appended)[ncol(label.appended)] <- c("Level")
label.appended
type.level.tibble <- label.appended %>% select(Type, Level)
type.level.tibble <- type.level.tibble %>% group_by(Level)

```



